---
title: "Survival modelling"
author: "Torgeir"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_notebook: default
  github_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
options(tibble.print_min = 5)
library(pander)     # for prettier outputs of summaries and tables etc.
# library(report)     # for reporting test-statistics             not available for this version of R
library(tidyverse)  # entering the tidyverse
library(tidymodels) # and bringing the model-toolbox
library(lubridate)  # lubricating the handling of dates
library(survival)   # the original survival-package
library(survminer)  # a plot-package for ggplot integration
library(vip)        # for variable importance plots
library(equatiomatic) # Model formula displayer in RMarkdown


obs <- readRDS("obs_surv_prepared.rds") # obs prepared for survival analysis (Time_to_Event_Torgeir5.R, line 133)
covs <- readRDS("CTloc_covs.rds")
```

Implementing the survival analysis into the tidymodels framework, as I like their approach of connecting different modelling-packages into one identical syntax.
<!-- Comments look like this, instead of the LaTeX % -->
------------------------------------------

# Setting up the data

```{r}
obs %>% 
  skimr::skim(validated_species, period) 
```


First, I'll filter out the species I will focus on. As the cameras were set up to detect lynx, I will filter out species that are significantly smaller. Squirrel and hare are examples of species that can't be expected to get detected every time they pass close to the expected travelling route, due to the average height of the cameras, and that they are not angled towards the ground.

In addition I filter out non specific groups (e.g. birds), infrequent sightings (i.e. < 50), and irrelevant groups (cattle, vehicles, humans). However, the "irrelevant" groups could be interesting to come back to later, to see if their presence have predictive power in my models.
```{r sp_focus, echo=TRUE}
fjern <- c("nothing","hund", "menneske", "kjoeretoey", "motorsykkel", "sykkel", "ukjent", 
           "sau", "ku", "fugl", "skogshons", "smagnagere", "andre_maardyr", "andre_pattedyr") # uninteresting or too general groups
passes <- obs %>% group_by(validated_species) %>% 
  summarise(count = n(),   # flashed = mean(flash, na.rm = T), # don't know if i can find a relevant use of this
            period = period, flash = flash) %>% 
  filter(!is.na(validated_species), !(validated_species %in% fjern))
  ggplot(passes) +
  geom_bar(aes(reorder(validated_species, count, FUN = mean)), position = "dodge") +  # reorders by mean count
  geom_hline(yintercept = 50) + coord_flip() # flip the axes
# removing small mammals
small <- c("maar", "ekorn", "hare")
p_sp_focus <- passes %>% 
  filter(count > 50, !validated_species  %in% small) %>%  
  ggplot(aes(reorder(validated_species, count, FUN = mean))) + coord_flip()
p_sp_focus + geom_bar(aes(fill = flash),position = "dodge") + geom_hline(yintercept = 50)
```
Having filtered out most sightings, we are left with the most common, large mammals. There are very few sightings of lynx when divided by times it was and wasn't flashed by a white LED. 
For the rest of the species in this plot, we can rest assured that we have a lot of datapoints, and that the species are large enough to be photo captured every time they pass the camera by the expected route.

## filtering out species
I don't know if this is specially useful for survival modelling, selecting variables and all that.
My models should only include one species per time anyway.
```{r obs_focus}
sp_focus = c("raadyr","rev","grevling","elg","hjort", "gaupe")
obs_focus <- obs %>% 
  filter(validated_species  %in%  sp_focus) %>% # filtering for our species
#  mutate() # mutating date col
  select(loc, date, validated_species, period, flash, flashed, t.diff, event, # selecting our variables
         datetime, ID, timeserie_id) #%>% #and including some ID variables for later
#  na.omit() %>%     # Exclude missing data
  # For creating models, it is better to have qualitative columns
  # encoded as factors (instead of character strings)
 # mutate_if(is.character, as.factor)
obs %>% 
  count(flash) %>% 
  mutate(prop = n/sum(n)) %>% 
  pander()
obs_focus %>% 
  count(flash, flashed) %>% 
  mutate(prop = n/sum(n)) %>% 
  pander()
```
Anyway, it's nice to see that the proportions of flash 1/0 stays the same after filtering out the focal species. This is probably mostly due to the control cameras.

Somewhat strange to consider the count of flashed during a flash-period, though.
I need to get a new grasp of what the flashed-column means, ie. how it's actually made.




# Build the model

### Baseline hazard function

The baseline hazard function $H_0(t)$ is the hazard at time $t$ when all predictors equals to zero.
You can think about the $H_0(t)$ being the intercept, although this is not strictly true. 
That is because in reality, $H_0(t)$ varies over time, and we can never know its true value.
However, we can calculate the hazard ratio between groups. We will never know a true hazard for any given group at a given time,
because we lack the knowledge of the shifting baseline, but we can know the relative hazard between groups.

In other words, for my case, I do not know the true probability (hazard) for any given species to get photo-captured by any of the
cameras in my study. Nor can I find it out, as this truly is an ever changing value. Still, I can measure the relative difference
in frequencies between the groups equipped with a white LED flash, and the groups only equipped with IR flash.

Ratios must stay constant over time, ie. if group B has twice the hazard of group A at time $t$, the assumption is that this relationship will stay that way at time $t + 1$. 
In other words, as I assume different species will have different reactions to the white LED, I cannot estimate the Hazard ratio between them.  _Or can I?_
Actually, I think I can't, because if a badger is indifferent to the white LED, his ratio to any other _reacting_ species will be constantly changing.

```{r mod0_raadyr}
sp = "raadyr"
obs_sp <- obs %>% filter(validated_species %in% sp & !period == "Control") # filtering out control + keeping only sp
mod0 <- coxph(Surv(t.diff, event, type = "right") ~ flashed, 
  data = obs_sp)                                            #  cleaner code with a temporary obs_sp
summary(mod0) 
```
So what am I really testing when I'm doing a survival analysis?
I'm testing whether or not the survival curves are significantly different from each other.
If it is, the summary-output of a model will show a p-value ( Pr(>|z|) ) lower than 0.05, and the lower and upper .95 interval should not contain the value _**1**_.

The mod0 for roe deer has a $Pr(>|z|) = 0.0444$ and the .95 confidence interval is $1.003$ to $1.309$.
In other words the model is right on the edge of being significant.
This seems to hint at an actual effect of the flash, but the p-score should keep us critical.


```{r}
# Survival probabilty against time.
fit <- survfit(Surv(t.diff, event, type = "right") ~ flashed, 
  data = obs_sp)
ggsurvplot(fit, data = obs_sp, title = "roe deer", 
          risk.table = T, break.time.by = 20, ggtheme = theme_minimal(),
          censor.shape = "|", censor.size = 3,
          tables.height = 0.2, tables.theme = theme_cleantable())
ggsurvplot(fit, data = obs_sp, title = "roe deer",
          risk.table = T, break.time.by = 10, ggtheme = theme_minimal(),
          conf.int = T, xlim = c(0,50), censor.shape = "|", censor.size = 3,
          tables.height = 0.2, tables.theme = theme_cleantable()
          #cumcensor = T
)
ggsurvplot(fit, data = obs_sp, title = "roe deer",
         coord_trans(x="log2"))

# A nicer way to visualize the coffecient estimate and the
# confidence interval. Since the interval doesn't overlap 
# 1 (hazard ratio of 1) the coeffecient estimate is significant.
ggforest(mod0, data = obs_sp)

# Diagnostics - Are we violating the proportional hazard
# assumption?

# Test the proportional hazards assumption
d.mod0 <- cox.zph(mod0)  
d.mod0 # Non-significant --> we can assume proportional hazards.
summary(fit)$table

# Can also look at Schoenfeld residuals, there should be no
# pattern with time
ggcoxzph(d.mod0)

```
When the confidence intervals are included in the plot, the white LED CI overlaps the IR curve almost constantly.
Therefore it doesn't seem to be a large effect of our flash.
Additionally, it is interesting to see that the effect is of the blits is luring, rather than scaring.
The survival time is shorter for white LED, which means that roe deer either gets drawn to the camera, or at least that their detection rate goes up.

The curves seem to cross eachother at a couple of points, but the Schoenfeld Individual test is non-significant,
so we can assume proportional hazards




-----------------------------------------------

After having compared the two hazards, ie. extracted a hazard ratio, I should check for confounding factors.
To do that, I add variables to the model formula, and check whether the intercept and the standard error of my group changes.
If they are approximately identical, the new factors don't seem to have any explanatory power.

Then, enter the _likelihood ratio test_. I'll use it to test if the full model is significantly better than the reduced model
$H_0$: no difference between models
$H_1$: the full model explains more (larger predictive power)
$^Gstat$ follows a $x^2$-distribution where df = k
P>|chi|:  if significant (<0.05), then $H_1$ is our model






------------------------------------------------------------


# In case I want to make a machine-learned model later
```{r training_split}
set.seed(121)
splits      <- obs %>% filter(!period == "Control") %>% # filter out control data
  initial_split(strata = flash)

obs_other <- training(splits)
obs_test  <- testing(splits)

# training set proportions by flash
obs_other %>% 
  count(flash) %>% 
  mutate(prop = n/sum(n)) %>% 
  pander()

# test set proportions by flash
obs_test  %>% 
  count(flash) %>% 
  mutate(prop = n/sum(n)) %>% 
  pander()
```
Inital split successfull
```{r}
set.seed(254)
val_set <- validation_split(obs_other, 
                            strata = flash, 
                            prop = 0.80)
val_set
```


-------------------------------------------------------------


# Session Info
```{r sessionInfo}
sessionInfo()
# packrat
# checkpoint
```

If you want your code to be reproducible in the long-run (i.e. so you can come back to run it next month or next year), you’ll need to track the versions of the packages that your code uses.
A rigorous approach is to use _packrat_, [link](http://rstudio.github.io/packrat/), which stores packages in your project directory,
or _checkpoint_, [link](https://github.com/RevolutionAnalytics/checkpoint), which will reinstall packages available on a specified date. A quick and dirty hack is to include a chunk that runs sessionInfo() — that won’t let you easily recreate your packages as they are today, but at least you’ll know what they were.