\thispagestyle{plain}
\begin{center}
	\Large
	\textbf{Flashing Large Mammals}
	
	\vspace{0.4cm}
	\large
	%Quantifying the effect of white LED flash on camera trapping detection rates
	Does white LED flashes in camera traps \\
	affect detection rates of target species?
	
	\vspace{0.4cm}
	\textbf{Torgeir Holmgard Valle}
	
	\vspace{0.9cm}
\end{center}
\newpage

\section{Abstract}

Camera trapping is a rapidly improving survey method that is largely targeted towards large mammals, and especially large carnivores.
Nonetheless, the cameras are triggered by all large and medium-sized animal species in the area, and thus gathers valuable data on the whole ecological community.
% And
White light flashes are sometimes utilized to get more detailed photos allowing for capture-recapture based population estimates for naturally marked species, like the Eurasian lynx (\textit{Lynx lynx}).

% But
However, the white light could function as a stressor or attractant for different species, which would affect density estimates. 
There are evidence of behavioural change in several mammal species, when exposed to a white flash, but quantifications on the detection rate of species are still lacking.


% In this study
Therefore, I investigated whether introducing an additional white LED camera trap (CT) at established CT sites affected the detection rates of the most common wild mammal species in the area.

% I predicted 
I predicted that the detection rate of nocturnal and crepuscular species would be altered as a response to the white light stimuli, and that the extent of the effect would depend on the species' visual sensitivity. 

% What happened was
I found a minimal effect of white LED, suggesting short term avoidance, which did not lead to longer term avoidance of the sites.

% An explanation could be
An explanation could be...


\textit{Keywords:} 
%Animal behaviour; % Tja, eit klart resultat av åtferd, men eg bruker liten tid på det i oppgåva
camera trap;
camera trap shyness; % To the point!
monitoring bias; % To the point
night-time photography; % Det er vertfall då det burde ha noko å sei
diel activity; % Nå som eg har brukt så mykje tid på det
density estimation; % Hovudgrunnen til å undersøke detection rates


%When in the abstract of an article, authors conclude an effect is “statistically equivalent,” the abstract should also include the equivalence bounds that are used to draw this conclusion.
%TODO Frå TOST-artikkelen til Daniel Lakens (2017)