\thispagestyle{plain}
%\LARGE
%\textbf{Acknowledgements}
\section*{Acknowledgement}
Studying biology at the University of Oslo has seen me fulfilling some of my dreams, like living abroad, visiting the Amazon, and studying mammals. Finally, it all culminated in writing this thesis.
I always assumed that writing a long text would be straightforward. 
However, doing some research on the subject, I've concluded that it is not.
Nevertheless, it is a fruitful endeavor, which has taught me many things.

I am grateful to my supervisors Atle Mysterud, Neri Horntvedt Thorsen and Inger Maren Rivrud and John Odden, for their help and contributions.
Atle, thank you for being a huge help in writing the thesis, conditioning me with positive reinforcements to see me reach the finishing line.
Neri, thank you for the colossal job with the data wrangling and setting up my analyses. You were speaking in codes, but I learned to decipher them.
Inger Maren, thank you for your writing advice, and tidying up my messy manuscript during my writing finale.
John, thank you for giving me the opportunity to work on a project related to lynx, overflowing with hours in the field.

Thanks to all the people working on the camera trap projects at NINA, and especially thanks to Solveig and Nina for always answering the phone and helping me during smaller and larger emergencies. There were a few of those!

During several hundred hours of field work, I was lucky enough to bring some friends along the way. Thank you all for making my field working days that much better. In particular, I'm grateful to Olav, who joined me for several merry weeks. Two fools were better than one, when venturing forth into gamble and peril.

Finally, thanks to my parents --- for being good parents. You've kept me sane these last months.