---
title: "Copy earlier mod1-7 of roe deer"
output: html_notebook
---

# Models with spatial covariates

```{r covs, include=FALSE}
covs<-readRDS("CTloc_covs.rds")
class(covs)
covs<-as.data.frame(covs) # Chaning class to data.fram and not a sf data.frame
# names(covs)
obs<-merge(obs, covs, by.x="loc", by.y="LokalitetID", all.x=TRUE, all.y=FALSE)
names(obs)

# Checking if there is any NAs in the covariates
lapply(1:ncol(obs), function(x){any(is.na(obs[,x]))})
colnames(obs)[8:10] # Not any in the covariates, but in some species... (valid_sp, distance, num_animals)
```


```{r mod1}
# Fitting model with spatial covariates
obs_sp <- obs %>%  # remaking obs_sp, to include the new covariates
  filter(species %in% sp & !period == "Control")

# Example with distance to forest roads and houses
mod1<-coxph(Surv(t.diff, event, type="right") ~ flashed +
              house_d2 + forestroad_d2, data=obs_sp)
summary(mod1)
ggforest(mod1, data = obs_sp)
```

Just adding the house_d2, and forestroad_d2 covariates completely changed the verdict on the effect of flashing.
Closeness to house has a slightly, but significant, _negative_ predictive power,
and closeness to forest roads has a ever slightly, but significant, _positive_ predictive power
However, I mistrust their confidence intervals. They are weirdly narrow.

<!--Also, Neri started a sentence commenting on what the negative value of house_d2 signifies, but didn't complete it.
-->
Anyways, just thinking about my spatial covariates, and what they represent, the way they are used in the model right now, is probably not what I intend.
I don't expect an animal's reaction to be linear with the distance to a road or a house.
Rather, I expect it to be exponentially larger, the closer the animal gets, ie. the closer the camera's position is to a road or a house, as that is where I observe them.
Worded differently, I don't think the difference of 1km and 3km to the nearest house will affect a roe deer, 
as much as the difference of 50m and 500m will.
Therefore I need to log transform the spatial covariates, and in doing so I have to be careful with 0-values.
Many cameras are positioned on or directly next to a forest road, and log transforming these values would create infinite-values. 


## Transforming the variables, and looking at correlations


```{r transform}
obs$forestroad_d2_ln<-ifelse(obs$forestroad_d2>0,log(obs$forestroad_d2), 0)
obs$forestroad_d2_io <- ifelse(obs$forestroad_d2>10, 0, 1)
obs$house_d2_ln<-ifelse(obs$house_d2>0,log(obs$house_d2), 0)
# plotting distances
qplot(forestroad_d2_io, loc, data =obs)

library(reshape2) # for melt function
p_theme <- theme(
  panel.grid.major.x = element_blank(),panel.grid.minor.x = element_blank(),
  panel.grid.minor.y = element_blank(),axis.text.x = element_blank())   
  #theme(panel.background = element_rect(fill = "white", colour = "grey50"))

obs %>% melt(id = "loc", measure = c("forestroad_d2", "house_d2")) %>% 
  ggplot(aes(as.factor(loc), value)) + geom_point(aes(col=variable)) + 
  labs(title = "Distance") + p_theme 
obs %>% melt(id = "loc", measure = c("forestroad_d2_ln", "house_d2_ln")) %>% 
  ggplot(aes(as.factor(loc), value)) + geom_point(aes(col=variable)) + 
  geom_smooth(span = 0.7) + labs(title = "Log-transformed distance") +
  p_theme #+ theme(legend.position = "top")
ggplot(obs, aes(forestroad_d2, house_d2)) + geom_point() + 
  geom_smooth(method = "glm", formula = y ~ x, se = T) +
  labs(title = "Distance correlation")
  #geom_smooth(span =.5)
library(corrplot)
covs %>% select(!c(1,3,4)) %>% cor() %>% 
corrplot(type = "upper", method = "number")

names(covs)
```
The two variables are almost independent of eachother, although the correlation increases somewhat when both are log-transformed. They will likely be able to predict variation complimentary to each other
Also, they are both independent to my flashed-variable.

Remembering my own question as stated in the intro:

*In this study, I will attempt to quantify how the usage of white LED flash affects the detection rate of the most common large mammal species in the area and whether this effect correlates with other factors [such] as urbanisation.*

Fores troads are absolutely a factor of human interference or something like that, but is not exactly what I'm getting at with urbanisation. Also, forest roads are known to attract many species [citation?] and are therefore used actively in camera trapping studies, my own included. So, including closeness to forest roads can probably account for some "attraction" to camera sites, or rather, higher detection frequencies. Maybe a TRUE/FALSE factor for camera being on forest road or not would be better.

On the other hand, distance to house is closer to my urbanisation statement, and is somewhat more of what I had in mind when articulating it.
Specifically, I was thinking about sources to Artificial Light At Night (ALAN), as a possible predictor for how animals react to white LED flash.
Proximity to houses is a somewhat good proxy for that, but still, I'll be missing the ALAN from other types of infrastructure, such as illuminated public roads, and heavily trafficked roads in general.

In any case, if I do get a hold of an actual proximity to ALAN-covariate (or similar), I need to be mindful of whether it is the ALAN or something correlating with the ALAN that can be causal to this relationship.


```{r mod2}
sp = "raadyr"
obs_sp <- obs %>%  # remaking obs_sp, to include the new covariates
  filter(species %in% sp & !period == "Control")

mod2<-coxph(Surv(t.diff, event, type="right")~flashed+ house_d2_ln + forestroad_d2_io, data=obs_sp)
summary(mod2)
ggforest(mod2, data = obs_sp)

# Test the proportional hazards assumption
d.mod2<-cox.zph(mod2) # Non-significant --> we can assume proportional hazards.
d.mod2

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod2)

```

Closeness to house, when log-transformed, was in fact highly significant as a negative predictor.
Forest roads are just barely positive, and unsignificant. 
This could be because most cameras weren't situated on forest roads, and therefore it would predictively have no effect.
The roads aren't a good predictor of where there are many animals, but rather, the route many animals would choose to travel.
In addition, they are less obstructed by vegetation, and therefore it is easier to trigger the camera, and easier to spot the animal.
Yes/no transforming the variable would also have the added benefit of removing most correlation with the house distance-variable.

The CI of the flashed variable has shifted ever so slightly, now including a Hazard ratio of 1. This model also leaves flash accounting for a positive trend in survival, ie. more frequent redetection of roe deer, but the effect doesn't explain a significant amount of the variation left after considering the other variables.


```{r mod3}
# Setting up model to test if the effect of flash is dependent on distance to house
mod3<-coxph(Surv(t.diff, event, type="right")~flashed*house_d2_ln + forestroad_d2_ln, data=obs_sp)
summary(mod3)
ggforest(mod3, data = obs_sp)

# Test the proportional hazards assumption
d.mod3<-cox.zph(mod3) # Non-significant --> we can assume proportional hazards.
d.mod3

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod3,font.main = 12, ggtheme = theme_classic2()) #caption = "Caption goes here"
```
The Global Schoenfeld test has a dangerously low p-value ($p = 0.06$), but it is still insignificant, and we may assume - carefully - proportional hazards.
Looking at the plots for flashed and flashed:house_d2_ln there are some outliers surpassing the dashed line. However, the Schoenfeld Individual Tests are all insignificant.

<<<<<<< HEAD
Interesting to note, is that the flashed effect increases wildly in it's confidence interval, leaving the house_d2_ln variable seeming diminished in it's variation / CI.
Importantly, though, there is no significance in the interaction between flash and distance to house.

```{r mod4}
# Setting up model to test if the effect of flash is dependent on distance to house
mod4 <- coxph(Surv(t.diff, event, type="right") ~ flashed * house_d2 +
                forestroad_d2_ln, data=obs_sp)
summary(mod4)
ggforest(mod4, data = obs_sp)

# Test the proportional hazards assumption
d.mod4<-cox.zph(mod4) # Non-significant --> we can assume proportional hazards.
d.mod4

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod4,font.main = 12, ggtheme = theme_classic2()) #caption = "Caption goes here"
```
Reverting the house_d2 variable to it's natural form ends up violating the global Schoenfield test (p:0.0067 !)

```{r mod5}
# Setting up model to test if the effect of flash is dependent on distance to house
mod5 <- coxph(Surv(t.diff, event, type="right") ~ flashed * forestroad_d2_ln +
                 house_d2_ln, data=obs_sp)
summary(mod5)
ggforest(mod5, data = obs_sp)

# Test the proportional hazards assumption
d.mod5<-cox.zph(mod5) # Non-significant --> we can assume proportional hazards.
d.mod5

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod5,font.main = 12, ggtheme = theme_classic2()) #caption = "Caption goes here"
```
The same thing happens when checking for an interaction between the log-transformed distance to forestroad-variable. The interaction seems to violate the individual test, but the model all together completely fails the global test.
Interesting to note how the predicted effect of the flash suddenly shifts to being negative!

Let's see what happens using the normal distance to forestroad covariate:


```{r mod6}
# Setting up model to test if the effect of flash is dependent on distance to house
mod6 <- coxph(Surv(t.diff, event, type="right") ~ flashed * forestroad_d2 +
                 house_d2_ln, data=obs_sp)
summary(mod6)
ggforest(mod6, data = obs_sp)

# Test the proportional hazards assumption
d.mod6<-cox.zph(mod6) # Non-significant --> we can assume proportional hazards.
d.mod6

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod6,font.main = 12, ggtheme = theme_classic2()) #caption = "Caption goes here"
```
Now, the individual test for the forestroad_d2 variable fails. The interaction does not fail, although the global test fails again, just not that strongly.

Something does seem to happen when flash is considered in conjunction with a _"luring"_ or attractive variable. 
There is something strange going on with the distance to forestroad-variable, so maybe a variable for trail type at the camera station would be a better fit. Strangely, only 3 locations are considered to be on a forestroad, which surprises me. I remember there to be more. At least there are more of tractor roads, and none of the "claimed" forest roads are in any way highly trafficked.

One way to extract a better forest road-variable could be to generate it from the species variable  (eg. __species %in% c("kjoeretoy", "motorsykkel", etc)__). 


## Micro habitat covariate!

```{r micro-habitat}
habitat <- readRDS("habitat.rds")
# table of habitat types on locations in obs
habitat %>% 
filter(loc %in% obs$loc) %>% 
group_by(habitat) %>% summarise(n = n()) %>% arrange(desc(n))

# join habitat-col with obs-object
obs <- obs %>%
  left_join(habitat, by = "loc")
# filtering out species
sp = "raadyr"
obs_sp <- obs %>%
  filter(species %in% sp)
# Setting up model to test if the effect of flash is dependent on distance to house
mod7 <- coxph(Surv(t.diff, event, type="right") ~ flashed + habitat +
                 house_d2_ln, data=obs_sp)
summary(mod7)
ggforest(mod7, data = obs_sp)

# Test the proportional hazards assumption
d.mod6<-cox.zph(mod7) # Non-significant --> we can assume proportional hazards.
d.mod6

# Can also look at Schoenfeld residuals, there should be no pattern with time
ggcoxzph(d.mod6,font.main = 12, ggtheme = theme_classic2()) #caption = "Caption goes here"
```


# Testing the models
## AIC-test
The best-fit model according to AIC is the one that explains the greatest amount of variation using the fewest possible independent variables.

[some tips on aic-referring in method section](https://www.scribbr.com/statistics/akaike-information-criterion/)
```{r AIC_roe}

# Which of the four different models have the lowest AIC?
mod.sel<-AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6,mod7)
mod.sel$d.AIC<-mod.sel$AIC-min(mod.sel$AIC)
mod.sel[order(mod.sel$d.AIC),]

# mod1$formula
# What is your conclusion? 
```
The AIC score determines model 1 as the best model, excluding the two last models violating the assumption of proportional hazards.
Model 1 was the model with the raw distance to feature-covariates. However, the other models are close on the score level, and should not be dismissed completely.
Maybe different transformations could fit the data better? At least, it is interesting to note that mod3, that had interaction with house_d2 scored on level with mod2 without an interaction.
The score signifies that the model has a better fit, and that the better fit is strong enough to be worth the penalty for a more complicated model!

And, maybe the forest road covariate isn't the best predictor as only three cameras seem to be classified as on a forest road.

`r mod1$formula ;mod1$coefficients`



## Likelihood ratio test
Then, enter the _likelihood ratio test_. I'll use it to test if the full model is significantly better than the reduced model
$H_0$: no difference between models
$H_1$: the full model explains more (larger predictive power)
$^Gstat$ follows a $x^2$-distribution where df = k
P>|chi|:  if significant (<0.05), then $H_1$ is our model

------------------------------------------------------------------------------------------
# Model with spatial covariates
Which I've gone away from, because I learned about the random effect! (1|variable)



```{r transformations}
obs %>% mutate(flash = ifelse(period=="Control","Control group","Had periods with white LED")) %>% 
  group_by(habitat, loc, flash, house_d2_ln, cam_mod) %>% nest() %>% 
  ggplot(aes(habitat, house_d2_ln)) +
  geom_point(aes(col = cam_mod), position = position_dodge(width=0.2)) +
  theme(legend.position = "right", legend.title = element_blank()) +
  facet_wrap(~ flash, ncol = 1) + coord_flip()
# names(covs)
library(corrplot)
obs %>% select(Latitude:cam_mod) %>% select(!c(Longitude, habitat)) %>% 
  mutate(cam_mod=ifelse(cam_mod=="Browning", 1, 0)) %>%
  cor() %>% corrplot(type = "upper", method = "number")
names(obs)
obs %>% group_by(species) %>% 
  summarise(count = n(), period = period, flash = flash, cam_mod=cam_mod) %>% 
  filter(species %in% c("hund", "menneske", "kjoeretoey", "motorsykkel", "sykkel", "sau", "ku")) %>% 
  ggplot() + geom_bar(aes(reorder(species, count, FUN = mean),
                          fill = cam_mod), position = "dodge") +  # reorders by mean count
  coord_flip() + theme(legend.position = "top") # flip the axes

```

There are no clear correlation between house distance and type of microhabitat.
Comparing the control-group with the cameras that had periods with white LED, there are some factors with imbalances between the group. The control group has of yet not been included in the models.
Comparing browning-cameras with reconyx-cameras, there are appearant differences in habitat choice.

Browning cameras are mostly set up by the Hunter&Fisher-society of Viken (or similar), and seems to have been put up on more convenient places in general.
This is important to note, as this variable seems to be a predictor for easily accessible sites.
Therefore, it could be a predictor for regular human presence as well!

The last barplot confirms just that. Fair enough, there are more sheep and cattle at Reconyx-sites, but for all other categorioes here, there is a sound bias of more human presence at the browning sites.
Nearly twice as many human passings (correlated with dogs), and many times more vehicles.

Slope and Forestroad_d2 doesn't correlate with anything, publicroad_d2 correlates with house_d2 which I consider a better fit to my question.
Elevation correlates with Latitude, which is worth considering.
Camera model is also correlated with Latitude (ie. more Browning models towards north), but luckily elevation and cam_mod seem to be independent.




```{r multi-univariate}
covariates <- c("flashed", "cam_mod", "forestroad_d2","house_d2", "house_d2_ln",
                "field_d2", "field_d2_ln","elev", "elev_ln", "slope")
univ_formulas <- sapply(covariates,
                        function(x) as.formula(paste('Surv(t.diff, event)~', x)))
                        
univ_models <- lapply( univ_formulas, function(x){coxph(x, data = obs_sp)})

# Extract data 
univ_results <- lapply(univ_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coeficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })
res <- t(as.data.frame(univ_results, check.names = FALSE))
as.data.frame(res)
```

Multiple univariate analyses of some coefficients reveal that house, forestroad and elevation has a significant impact on time to event.
House_d2 has negative impact, and elevation has a barely negative one.
Forestroad has a barely positive one, but I've decided on a better replacement for this covariate, which is a categorisation of the micro habitat at each site.

I'll include all the significant variables except forestroad in the multivariate model, and add habitat. Also, I will only use the log-transformed counterparts in the house and elevation variables, as the normal scale predicts 1(1-1) as influence.




### Model 1 |  ~habitat + cam_mod + elev_ln + house_d2

```{r mod1}
sp="raadyr"
# Fitting model with spatial covariates
obs_sp <- obs %>%  # remaking obs_sp, to include the new covariates
  filter(species %in% sp & !period == "Control")

# Example with distance to forest roads and houses
mod1<-coxph(Surv(t.diff, event, type="right") ~ habitat + cam_mod + elev_ln +
              house_d2_ln + flashed, data=obs_sp)
summary(mod1)
ggforest(mod1, data = obs_sp)

# Test the proportional hazards assumption
d.mod1 <- cox.zph(mod1)  
d.mod1 # Non-significant --> we can assume proportional hazards.

# Can also look at Schoenfeld residuals, there should be no
# pattern with time
ggcoxzph(d.mod1, font.main = 12, ggtheme = theme_classic2())
```

Elevation is deemed non-significant, as have the sub-variate cliff. Although it could have been interesting to try out an interaction there, that seems overly complex to me.
Canyon is leaning towards having a postive effect on time to event, but is also considered non-significant.
Foot path, forest road, and wildlife trail are all highly significant, and positive predictors, maybe even more so after considering the variation accounted for in camera models (ie. browning cameras were biased towards foot paths and forest roads, as well as being passed by more humans and vehicles).

Camera model and distance to nearest house are both significant as well.
The negative contribution of distance to house is counter intuitive. The model tells us what higher values of a variable predicts for time to event. 
In this case, higher values for distance to house means lower probability of an event happening (ie. detection of species). 
>A positive sign means that the hazard (risk of death) is higher, and thus the prognosis worse, for subjects with higher values of that variable.

In other words, the farther away a site is from a house, the less likely it is to detect a roe deer. Which also means, the closer a site is to a house, the more likely the camera is of detecting a deer!
This is not the effect I was expecting when creating the model, and it isn't necessarily because of the houses themselves, but rather something correlating with these houses. _Namely, fields_.


Strangely, field_d2_ln has a low correlation with house_d2_ln, although they correlate with the normal variant of themselves and each other.
The transformation slightly alters both of them, resulting in a muddling of the correlation when both of them are transformed.
Still, as both covariates correlates in their natural form, that is their true relationship, and they should not be included simultaneously.

In the next models I will therefore remove non-significant whole covariates (e.g. elevation), and then make one variant with house_d2_ln and one with field_d2_ln.

I will not exclude any sub-variables of the habitat although they are considered non-significant. This is because removing some of them could ultimately alter differing sub-variates relation to the intercept-variate (boulder) and the different statistics are calculated out from all variates together, as mentioned in [this article](https://towardsdatascience.com/stopping-stepwise-why-stepwise-selection-is-bad-and-what-you-should-use-instead-90818b3f52df).




### Model 2 |  ~habitat + cam_mod + ( house_d2_ln ||  field_d2_ln )

```{r mod2-house}
library(coxme)
sp="raadyr"
# Fitting model with spatial covariates
obs_sp <- obs %>%  # remaking obs_sp, to include the new covariates
  filter(species %in% sp & !period == "Control")

# Example with distance to forest roads and houses
mod2a<-coxme(Surv(t.diff, event, type="right") ~  flashed + #house_d2_ln + # + cam_mod + house_d2_ln +
               (1|loc) + (1|habitat), data=obs_sp)
summary(mod2a)
#ggforest(mod2a, data = obs_sp)

ranef(mod2a)

exp(7.640597e-01)

# Test the proportional hazards assumption
d.mod2a <- cox.zph(mod2a)  
d.mod2a # Non-significant --> we can assume proportional hazards.

# Can also look at Schoenfeld residuals, there should be no
# pattern with time
ggcoxzph(d.mod2a, font.main = 12, ggtheme = theme_classic2())
```


```{r mod2-field}
sp="raadyr"
# Fitting model with spatial covariates
obs_sp <- obs %>%  # remaking obs_sp, to include the new covariates
  filter(species %in% sp & !period == "Control")

# Example with distance to forest roads and houses
mod2b<-coxph(Surv(t.diff, event, type="right") ~ habitat + cam_mod +
              field_d2_ln + (1|loc), data=obs_sp)
summary(mod2b)
ggforest(mod2b, data = obs_sp)

# Test the proportional hazards assumption
d.mod2b <- cox.zph(mod2b)  
d.mod2b # Non-significant --> we can assume proportional hazards.

# Can also look at Schoenfeld residuals, there should be no
# pattern with time
ggcoxzph(d.mod2b, font.main = 12, ggtheme = theme_classic2())
```

ever so slightly, lower AIC score for the field-model.


```{r survplot-for-complete-model}
# Plot the baseline survival function
ggsurvplot(survfit(mod1), color = "#2E9FDF",
           ggtheme = theme_minimal(), data = obs_sp)
```

```{r comparative-survplot}
# Create the new data  
flash_df <-
  with(obs_sp,
     data.frame(flashed = c(0, 1), 
                age = rep(mean(age, na.rm = TRUE), 2), #continuous - mean as mean
                ph.ecog = c(1, 1) # discrete - lowest value as mean
                )     )
flash_df
```


