---
title: "Flashing Large Mammals - exploring the data"
author: "Torgeir Holmgard Valle"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r setup}
library(tidyverse)
library(lubridate)
library(overlap)
library(sf) # for handling spatial data

# All the tibbles from Data_explor_setup.R, and package + function from Density.R
obs      <- readRDS("Observations_prepared1.rds") %>%
  mutate(species = validated_species)
stations <- readRDS("stations.rds")
# Tweaking obs to resemble time.dep in glmm_sp.Rmd
ctrl <- c("Control_1", "Control_2", "Control_3","Control_4")
effort   <- readRDS("Effort_prepared.rds") %>%
    mutate(flash = factor(
            ifelse(period %in% ctrl, "Control", flash)), # Including Control in flash
          week = lubridate::isoweek(date), # extracting week-column
          period = factor(period, labels = c("IR_1", "IR_2", "LED_1", "LED_2", # turning period into a factor
                          "Control_1", "Control_2", "Control_3", "Control_4"))) # and relabeling periods
effort <- effort %>% 
   mutate(flash = factor(flash, levels = c("Control","0","1"),
                ordered = T, labels = c("Control","IR","LED"))) # relevel to make Control the model intercept
```


# Presenting the data
```{r}
min(obs$date) # [1] "2019-01-15"
max(obs$date) # [1] "2020-02-17"  #but last field date was 2020-02-26 at camera 850 & 822!!

sp_focus <- c("ekorn", "elg", "grevling", "hare", "raadyr", "rev")
obstation <- obs %>% left_join(stations, by = "loc") %>% filter(species %in% sp_focus) %>%
  unite(abc, abc, flash) # unite lager ny faktor med alle kombinasjoner

#obstation %>% with(table(species, abc)) %>% 
 # knitr::kable(caption = "Table with kable")

# Total nr of image (rapidfire)series can be presented by:
length(unique(obs$timeserie_id)) # identical numbers for timeserie_id and image_id
length(unique(obs$image_id))     # is it because only one picture is selected per timeseries?
# obs_by_loc <- obs %>% group_by(loc) %>% summarise(n = length(timeserie_id)) %>% arrange(desc(n))
# plot(obs_by_loc)

x <- obs %>% left_join(stations) %>% group_by(abc, loc, cam_mod) %>% 
  summarise(n=n())
x %>% group_by(abc) %>% count() %>% pander::pander()
x %>% group_by(cam_mod) %>% count() %>% pander::pander()

stations %>% group_by(cam_mod) %>% count() %>% pander::pander()
```

## Timeseries plot of periods from the troubleshooting and organisation with Neri

```{r effort-facet-setup, echo=FALSE,message=FALSE}
c <- effort %>% filter(period %in% ctrl) %>% 
  group_by(period) %>% # period breaks of cam C
  summarise(c = unique(min(date))) # to use in top x-axis
visits <- read_csv("visits.csv") %>% filter(!loc==664) # error in date-column
field_work <- list()
field_work$v1 = ymd(range(visits$v_date1,na.rm = T)) # range of dates
field_work$v2 = ymd(range(visits$v_date2,na.rm = T)) # of each field work period
field_work$v3 = ymd(range(visits$v_date3,na.rm = T))
field_work$v4 = ymd(range(visits$v_date4,na.rm = T)) # ymd() from lubridate
field_work$v5 = ymd(range(visits$v_date5,na.rm = T)) # similar to as.Date()
field_work <- as_tibble(field_work)
# vector to make break-points for geom_vline and bottom x-axis
f_work <- c(field_work$v1,field_work$v2,field_work$v3,field_work$v4,field_work$v5)
# enter values later used in geom_rect to shade field work periods
rects <- data.frame(xstart = f_work[c(1,3,5,7,9)],
                    xend  = f_work[c(2,4,6,8,10)], field = "work")
```


```{r effort-facet}
p_eff <- effort %>%
  mutate(trt_gr = ifelse(period %in% ctrl, "Control", "Treatment"), # faceting factor
         # renaming control-groups in order to extract period-breaks, and not confuse with ctrl_1,2,3,4
         flash = factor(ifelse(period %in% ctrl[c(2,4)], 4, flash), ordered = T,
                      levels = c(1,4,2,3),
                      labels = c("Control", "period splits added during analysis","IR", "White LED"))) %>% 
ggplot() +
  facet_grid(rows = "trt_gr", scales = "free_y", space = "free_y") +
  geom_point(aes(date, as.factor(loc), col=flash), size = 0.9) +
  theme_classic() + labs(y = "Location") +
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        legend.position = "top")+
  scale_color_manual(values = c("#fdae61","#f46d43",     # ctrl-col
                                 "#abd9e9","#4575b4")) + # trt-col
  scale_x_date(NULL, breaks = f_work, date_labels = "%d-%b",
               sec.axis = sec_axis(~., breaks = c(c$c,f_work[10]), labels = scales::date_format("%d-%b %y") ) ) +
  geom_vline(xintercept = f_work, linetype = "dashed",  alpha =.5) +
  geom_rect(data = rects, aes(xmin = xstart, xmax = xend, ymin = -Inf, ymax = Inf),
            alpha = 0.1 )
p_eff +   theme(legend.title = element_blank()) +
  guides(colour = guide_legend(override.aes = list(size = 2), ncol = 2, 
                               reverse = F, byrow = T), #adjust rows of legend
                x = guide_axis(n.dodge = 2)) # dodge axis text into n rows
                # x = guide_axis(angle = -10))  #adjust axis-text angle
```

 

-----------------------------

# Some final plots I want to add in my thesis


## Maps

```{r mapsetup, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE}
library(sf)

Viken <- readRDS("viken.rds") # name layer from NVE municipal borders.gfs
str(Viken)

loc_sfg <- st_multipoint(cbind(stations$x,stations$y)) 
loc_sfc <- st_sfc(loc_sfg, crs = 23032) #23032 for utm zone 32 (23033 for utm33)
# Create data.frame with attributes
loc_df <- data.frame(loc = stations$loc,
                        cam_mod = stations$cam_mod,
                        abc = stations$abc)
# Combine data.frame and spatial data
loc_sf <- st_sf(loc_df, geometry = loc_sfc)
loc_sf
class(loc_sf)
str(loc_sf)
# loc_sf <- mutate(loc_sf, 
#                    x = purrr::map_dbl(geometry, 1), 
#                    y = purrr::map_dbl(geometry, 2))

```
```{r maps-scandi}
library(maps)
library(tmap)
Norge <- map("world",'Norway', plot = FALSE, fill = TRUE)
Norge <- st_as_sf(Norge)
ggplot() + geom_sf(data = Norge)
Scandinavia <- st_as_sf(map("world",c("Norway","Sweden", "Finland","Denmark","UK",
                                      "Germany", "Estonia", "Latvia","Lithuania"), plot = F, fill = T))
# loc_trans <- st_sfc(loc_sf, crs=)
xy <- st_bbox(loc_sf) # retrieve xlim+ylim from geometry
ggplot() +
  geom_sf(data =Scandinavia) +
#  geom_sf(data = Viken, fill = "blue", alpha = .02) +  # Viken polygon
#  geom_sf(data = loc_sf, fill = "blue", alpha = .2) +  # loc range

  # annotate("rect", xmin = xy[1], xmax = xy[3],
  #        ymin = xy[2], ymax = xy[4],
  #        alpha = .4) +
   annotate(geom="rect", xmin = 9.3, xmax = 11.5, # i samsvar med excel-
            ymin = 59.3, ymax = 60.5,              # kjørearket mitt
            alpha = .5, col = "black") +
  coord_sf(xlim = c(0.7,25), ylim = c(57,71)) +
  theme_bw()

# ggplot(Viken) +
#   stat_sf_coordinates()
```

```{r}
# library(tmap)
# tmap_mode("plot")
# tm_basemap() + 
#   tm_polygons(Viken) + #will only draw polygon, not additional layers
#   tm_shape(loc_sf) + tm_bubbles(col = "red")


# Library
library(ggmap)
#register_google(key = "your code")
library(gridExtra)
# For stamen map, you have to give the extremity of the window you are looking at. here is an example with the watercolor background (Around brisbane)
Norge1 <- get_stamenmap( bbox = c(left = 0, bottom = 57, right = 25, top = 71), zoom = 4)
Norge2 <- get_stamenmap( bbox = c(left = 0, bottom = 57, right = 25, top = 71), zoom = 4, maptype = "toner")
Viken1 <- get_stamenmap( bbox = c(left = 9.28, bottom = 59.2, right = 11.4, top =  60.5), zoom = 4, maptype = "toner")


ggmap(Norge1) + 
  theme_void() +
  theme(
    plot.title = element_text(colour = "orange"), 
    panel.border = element_rect(colour = "grey", fill=NA, size=2)
  ) +   annotate(geom="rect", xmin = 9.28, xmax = 11.4, # i samsvar med excel-
            ymin = 59.2, ymax = 60.5,              # kjørearket mitt
            alpha = .5, col = "black")

```



```{r maps}
loc_brown <- filter(loc_sf, cam_mod %in% "Browning")
ggplot(data = Viken) +
        geom_sf() +     # Viken
        # geom_sf_label(aes(label = navn)) + #denne kommandoen kan også hente ut namnelapper
       # geom_sf_text(aes(label = navn)) + #denne kommandoen kan også hente ut namnelapper
        geom_sf(data = loc_brown, size = 2,   # CT sites
        show.legend = "point", color="brown") + #alle loc får begge cam_mod levels ...
        theme_minimal() +
        theme(axis.title = element_blank()) +
        coord_sf(xlim = xy[c(1,3)], ylim = xy[c(2,4)])

```













## Which species did we get?
```{r sp_focus, echo=TRUE}
sp_all <- c("hare", "elg", "rev", "grevling", "maar", "gaupe", "ekorn", "raadyr", "hjort") # interesting species with enough datapoints
by_sp <- obs %>%
  left_join(stations, by = "loc") %>%
  group_by(species)
passes <- summarise(by_sp,
  count = n(),   # flashed = mean(flash, na.rm = T), # don't know if i can find a relevant use of this
  abc = abc, period = period, flash = flash)
passes %>% 
  filter(!is.na(species), !(species %in% "nothing")) %>% 
  ggplot() +
  geom_bar(aes(reorder(species, count, FUN = mean)), position = "dodge") +  # shows actual counts of each species
  geom_hline(yintercept = 200) + coord_flip() # flip the axes

# most datapoints on ekorn, elg, grevling, hare, raadyr, rev, 
fjern <- c("nothing","hund", "menneske", "kjoeretoey", "motorsykkel", "ukjent", "sau", "ku", "fugl") # uninteresting or too general groups
p_sp_focus <- passes %>% 
  filter(count > 200, !(species %in% fjern), !is.na(species)) %>%  # removing low counts and sp in 'fjern'
  ggplot() + scale_fill_brewer() + geom_hline(yintercept = 200) + coord_flip()
p_sp_focus + geom_bar(aes(reorder(species, count, FUN = mean)), position = "dodge") + geom_hline(yintercept = 200) # same as first plot, more filtered
p_sp_focus + geom_bar(aes(abc, fill = species), position = "dodge")  # same as first plot, more filtered
p_sp_focus + geom_bar(aes(period, fill = species), position = "dodge")
p_sp_focus + geom_bar(aes(flash, fill = species), position = "dodge")
```

Removing sightings of "nothing", things related to humans, as well as NAs and birds (too general group), I am left with quite a few species still. Filtering for counts lower than 200 renders me the selection shown in the dark themed plots.

Roe deer and red fox are the species with decidedly most data. The flash true/false-plot reveals similar proportions of detection abundances between the eight most common species.

## Density plots showing activity patterns with and without flash



```{r overlap-func, echo=FALSE}
# functions building on overlap package
overlap_flash <- function(x, colour = c("black", "blue"), rug = FALSE, main = str_to_title(x), ...) {
  x0 <- obs[obs$species %in% x & obs$flash == FALSE, ]$rad
  x1 <- obs[obs$species %in% x & obs$flash == TRUE, ]$rad
  overlapPlot(x1, x0, linecol = rep(colour, 2), rug = rug, main = main, 
              linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey", ...)
  x.est <- round(overlapEst(x1, x0, type = "Dhat4"), 2)
  n0 <- length(x0)
  n1 <- length(x1)
  legend("top", legend = str_c("Dhat4", x.est, sep = " = "), bty = "n")
#  legend("bottomleft", legend = c(str_c("LED n", n1, sep = " = "), str_c("    IR n", n0, sep = " = ")),
#         col = colour, lty = c(1, 2), lwd = 2, bty = "7")
}
overlap_flash_2sp <- function(x, y, col_x, col_y, rug = FALSE) {
  x0 <- obs[obs$species %in% x & obs$flash == FALSE, ]$rad # - - | | - -    = FALSE
  x1 <- obs[obs$species %in% x & obs$flash == TRUE, ]$rad  # sp X when flash = TRUE
  y0 <- obs[obs$species %in% y & obs$flash == FALSE, ]$rad # - - | | - -    = FALSE
  y1 <- obs[obs$species %in% y & obs$flash == TRUE, ]$rad # sp Y when flash = TRUE
  par(mfrow = c(2, 2)) # setting plot parameters: 2x2 windows
  overlap_flash(x, col_x, rug) # using the overlap_flash function for single species w/w/o flash-plot
  overlap_flash(y, col_y, rug) #                 - - | | - -
  overlapPlot(x0, y0,  # overlapping 2 species plot - flash = FALSE
    main = "IR", rug = rug, linecol = c(col_x, col_y),
    linetype = c(2, 2), linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey"
  )
  xy0.est <- round(overlapEst(x0, y0, type = "Dhat4"), 3)  # estimated overlap using Dhat4 - flash = FALSE
  xy1.est <- round(overlapEst(x1, y1, type = "Dhat4"), 3)  #           - - | | - -         - flash = TRUE
  legend("top", legend = str_c("Dhat4", xy0.est, sep = " = "), bty = "n") # Dhat4 label
  legend("bottomleft", legend = c(x, y), col = c(col_x, col_y), lty = c(2, 2), lwd = 2, bty = "n") # legend
  overlapPlot(x1, y1,  # overlapping 2 species plot - flash = TRUE
    main = "white LED", rug = rug, linecol = c(col_x, col_y),
    linetype = c(1, 1), linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey"
  )
  legend("top", legend = str_c("Dhat4", xy1.est, sep = " = "), bty = "n")
  legend("bottomleft", legend = c(x, y), col = c(col_x, col_y), lty = c(1, 1), lwd = 2, bty = "n")
  par(mfrow = c(1, 1)) # return normal plot parameters: 1x1 windows
}
obs$hour <- as.numeric(format(obs$datetime, "%H")) # setting up time cols 
obs$mins <- as.numeric(format(obs$datetime, "%M")) # in a way that overlap
obs$rad <- ((obs$hour * 60 + obs$mins)/(24 * 60)) * 2 * pi  # understands
```



```{r overlap-flash}
overlap_flash("rev", "red", rug = TRUE) 
overlap_flash_2sp("rev", "raadyr", "red", "blue", rug=TRUE)
```

Sites without white flash produces a more bumpy curve than does the sites with a flash. Proportion of foxes at sites are markedly lower before sunrise, and then higher afterwards. Could this simply be because of a lower detection rate when lacking the additional white flash? There is a resembling phenomenon happening in the evening twilight as well, right before the peak activity time of the fox, which happens before midnight.

The Dhat4 calculation reveals a larger difference in activity for foxes, than for roe deer, but seems to mainly stem from the twilight hours. Thus, it could be because of the size of the animals, rather than a reaction to the flash.

### Separate plot for each species

```{r, eval=FALSE}
sp <- c("rev", "raadyr", "grevling", "elg", "hjort", "gaupe", "ekorn", "hare", "maar")
# 1. Open jpeg file
png("png/rev.png")
# jpeg(cat(".jpg"), width = 350, height = "350")
# 2. Create the plot
overlap_flash(sp[1], rug = TRUE, main = "")
# 3. Close the file
dev.off()
```


### One plot for all species

```{r overlap-flash-sp}

sp <- c("rev", "raadyr", "grevling", "elg", "hjort", "gaupe", "ekorn", "hare", "maar")
par(mfrow = c(3, 3))
 par(cex = 0.6)
 par(mar = c(0, 0, 0, 0), oma = c(4, 4, 0.5, 0.5))
 par(tcl = -0.25)
 par(mgp = c(2, 0.6, 0))
overlap_flash(sp[1], main=NULL, rug=T) 
   legend("top", legend = str_c("\n ",sp[1]), bty = "n") # art 
for (i in sp[2:3]) {
 overlap_flash(i, axes = FALSE,  ann=F, rug=T, xaxt="n")
  legend("top", legend = str_c("\n ",i), bty = "n") # art
}
overlap_flash(sp[4], main=NULL, rug=T) 
   legend("top", legend = str_c("\n ",sp[4]), bty = "n") # art 
   legend(x=-4.2, y=0.018, legend = c("IR", "white LED"), #x+y sets legend position
         col = c("black","blue"), lty = c(1, 2), lwd = 2, bty = "o")
for (i in sp[5:6]){
 overlap_flash(i, axes = FALSE,  ann=F, rug=T) 
   legend("top", legend = str_c("\n ",i), bty = "n") # art
}
overlap_flash(sp[7], main=NULL, rug=T) 
   legend("top", legend = str_c("\n ",sp[7]), bty = "n") # art 
for (i in sp[8:9]){
 overlap_flash(i, axes = FALSE,  ann=F, rug=T) 
   legend("top", legend = str_c("\n ",i), bty = "n") # art
}
 mtext("Time", side = 1, outer = TRUE, cex = 0.7, line = 2.2,
 col = "grey20")
 mtext("Density", side = 2, outer = TRUE, cex = 0.7, line = 2.2,
 col = "grey20")
```



### Controlling for day-length
```{r suntime}
# Dates <- as.POSIXct(obs$date, tz="CET")
# coords <- matrix(c(60.1, 10.6), nrow=1) # Rett nord for Oslo i longlat
# Coords <- sp::SpatialPoints(coords, proj4string=sp::CRS("+proj=longlat +datum=WGS84"))
# st <- sunTime(obs$rad, Dates, Coords)
# 
# par(mfrow=2:1)
# densityPlot(st, col='red', lwd=2, xaxt='n', main="Sun time")
# axis(1, at=c(0, 6, 12, 18, 24),
#   labels=c("midnight", "sunrise", "noon", "sunset", "midnight"))
# densityPlot(simCalls$time, lwd = 2, main = "Clock time")
# par(mfrow=c(1,1))

# TODO # Sjå density.r for oppdatert skript 
```



## Box plots
### of periods
```{r box_plots_periods}
freq     <- readRDS("freq.rds") %>%  #from line 79 in Data_exploration2_nesting.R
  mutate(species = validated_species)
# Setting 20 days as an arbitrary limit
freq <- freq[freq$n.days > 19, ]
# frequency on periods, single or plural species
sp <- c("elg", "grevling","hjort", "maar", "raadyr", "rev") # Species with the most datapoints
p_freq_per <- freq %>% # plot of frequency grouped by period
  filter(species %in% sp) %>%
  left_join(stations, by = "loc") %>%
  ggplot(aes(period, freq)) +
  facet_wrap(~species, scales = "free", nrow = 2) +
  labs(title = "Frequencies of species in periods")

# coloured with abc-grouping
p_freq_per + geom_boxplot(outlier.shape = NA) + # Remove outliers when overlaying boxplot with original data points
  geom_jitter(aes(col = abc), width = 0.1) + coord_flip() # add points for each camera

```

### of flash on/of
```{r, box_abc_flash}
freq %>% # plot of frequency grouped by abc and flash T/F
  filter(species %in% sp) %>%
  left_join(stations, by = "loc") %>%
  ggplot(aes(abc, freq)) +
  facet_wrap(~species, scales = "free", nrow = 2) + # scales = "free"
  labs(title = "Frequencies of species w/ | w/o flash") +
  geom_boxplot(aes(fill = flash))
```











### Blank images
```{r blanks, eval = FALSE}
# Filtering out timelapse-pictures in Reconyx (455 is the only Browning-camera that has significant number of pictures here)
timelapse <- obs %>% left_join(stations, by = "loc") %>%
  filter(hour %in% c(8, 13) & mins == 0  &  secs == 0  &  species == "nothing") # == "nothing" to single out tinelapse-photos
summary(timelapse)
table(timelapse$hour[timelapse$cam_mod =="Reconyx"])

table(timelapse$cam_mod)


# table(timelapse$loc[timelapse$cam_mod == "Browning"])  #455 maybe has two cameras?
# table(timelapse$species)  # when val_sp == "nothing" not specified: 10 other species present
# table(timelapse$species[timelapse$hour == 13]) # elg:1, maar:2, menneske:3, raadyr:1
# table(timelapse$species[timelapse$hour == 8]) # ekorn:1, fugl:9, gaupe:1, katt:1, menneske:1, raadyr:4, rev:2, ukjent:4

# Differentiate "nothing"-sequences and "nothing"-one-shots
# species == "nothing" #by loc,
# and making an "independent event" criteria to distinguish repeated vegetation-triggering and other single-triggered events. 
# TODO
library(data.table)
setDT(obs)[, event_id := 1L + cumsum(c(0L, diff(datetime) > 720)), by=.(loc, species)]

# TODO obs %>% group_by(species, event_id) %>% filter(species == "nothing") %>% table(obs$loc)


names(obs)
```




# Session Info
```{r sessionInfo}
sessionInfo()
# packrat
# checkpoint
```

If you want your code to be reproducible in the long-run (i.e. so you can come back to run it next month or next year), you’ll need to track the versions of the packages that your code uses.
A rigorous approach is to use _packrat_, [link](http://rstudio.github.io/packrat/), which stores packages in your project directory,
or _checkpoint_, [link](https://github.com/RevolutionAnalytics/checkpoint), which will reinstall packages available on a specified date. A quick and dirty hack is to include a chunk that runs sessionInfo() — that won’t let you easily recreate your packages as they are today, but at least you’ll know what they were.