---
title: "Flashing Large Mammals - exploring the data"
author: "Torgeir"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: github_document
---

```{r setup}
library(tidyverse)
library(lubridate)
library(overlap)
# All the tibbles from Data_explor_setup.R, and package + function from Density.R
  effort   <- readRDS("Effort_prepared.rds")
  obs      <- readRDS("Observations_prepared1.rds")
  stations <- readRDS("stations.rds")
```

# Presenting the data
```{r}
min(obs$date) # [1] "2019-01-15"
max(obs$date) # [1] "2020-02-17"  #but last field date was 2020-02-26 at camera 850 & 822!!

sp_focus <- c("ekorn", "elg", "grevling", "hare", "raadyr", "rev")
obstation <- obs %>% left_join(stations, by = "loc") %>% filter(validated_species %in% sp_focus) %>% unite(abc, abc, flash) # unite lager ny faktor med alle kombinasjoner

#obstation %>% with(table(validated_species, abc)) %>% 
 # knitr::kable(caption = "Table with kable")

# Total nr of image (rapidfire)series can be presented by:
length(unique(obs$timeserie_id)) # identical numbers for timeserie_id and image_id
length(unique(obs$image_id))     # is it because only one picture is selected per timeseries?
obs_by_loc <- obs %>% group_by(loc) %>% summarise(n = length(timeserie_id)) %>% arrange(desc(n))
# plot(obs_by_loc)
```

## Timeseries plot of periods from the troubleshooting and organisation with Neri

```{r effort-facet}
ctrl <- c("Control_1","Control_2","Control_3","Control_4")
effort %>% filter(!period %in% ctrl) %>% 
ggplot()+
  geom_point(aes(date, as.factor(loc), col=factor(flash))) +
  theme_classic() + labs(y = "Location") +
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        legend.position = "top")
effort %>% filter(period %in% ctrl) %>% 
ggplot()+
  geom_point(aes(date, as.factor(loc), col = period)) +
  theme_classic() + labs(y = "Location") +
  theme(axis.ticks.y = element_blank(), axis.text.y = element_blank(),
        legend.position = "top")
#  theme(axis.text.x = element_text(angle = 20))   # angle endrer tekst, for å unngå overlapp
#  scale_y_discrete(guide = guide_axis(check.overlap = T)) # tar seg av overlappande tekst
```



## Which species did we get?
```{r sp_focus, echo=TRUE}
sp_all <- c("hare", "elg", "rev", "grevling", "maar", "gaupe", "ekorn", "raadyr", "hjort") # interesting species with enough datapoints
by_sp <- obs %>%
  left_join(stations, by = "loc") %>%
  group_by(validated_species)
passes <- summarise(by_sp,
  count = n(),   # flashed = mean(flash, na.rm = T), # don't know if i can find a relevant use of this
  abc = abc, period = period, flash = flash)
passes %>% 
  filter(!is.na(validated_species), !(validated_species %in% "nothing")) %>% 
  ggplot() +
  geom_bar(aes(reorder(validated_species, count, FUN = mean)), position = "dodge") +  # shows actual counts of each species
  geom_hline(yintercept = 200) + coord_flip() # flip the axes

# most datapoints on ekorn, elg, grevling, hare, raadyr, rev, 
fjern <- c("nothing","hund", "menneske", "kjoeretoey", "motorsykkel", "ukjent", "sau", "ku", "fugl") # uninteresting or too general groups
p_sp_focus <- passes %>% 
  filter(count > 200, !(validated_species %in% fjern), !is.na(validated_species)) %>%  # removing low counts and sp in 'fjern'
  ggplot() + scale_fill_brewer() + geom_hline(yintercept = 200) + coord_flip()
p_sp_focus + geom_bar(aes(reorder(validated_species, count, FUN = mean)), position = "dodge") + geom_hline(yintercept = 200) # same as first plot, more filtered
p_sp_focus + geom_bar(aes(abc, fill = validated_species), position = "dodge")  # same as first plot, more filtered
p_sp_focus + geom_bar(aes(period, fill = validated_species), position = "dodge")
p_sp_focus + geom_bar(aes(flash, fill = validated_species), position = "dodge")
```

Removing sightings of "nothing", things related to humans, as well as NAs and birds (too general group), I am left with quite a few species still. Filtering for counts lower than 200 renders me the selection shown in the dark themed plots.

Roe deer and red fox are the species with decidedly most data. The flash true/false-plot reveals similar proportions of detection abundances between the eight most common species.

## Density plots showing activity patterns with and without flash



```{r overlap-func, echo=FALSE}
# functions building on overlap package
overlap_flash <- function(x, colour = c("black", "blue"), rug = FALSE, main = str_to_title(x), ...) {
  x0 <- obs[obs$validated_species %in% x & obs$flash == FALSE, ]$rad
  x1 <- obs[obs$validated_species %in% x & obs$flash == TRUE, ]$rad
  overlapPlot(x1, x0, linecol = rep(colour, 2), rug = rug, main = main, 
              linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey", ...)
  x.est <- round(overlapEst(x1, x0, type = "Dhat4"), 2)
  n0 <- length(x0)
  n1 <- length(x1)
  legend("top", legend = str_c("Dhat4", x.est, sep = " = "), bty = "n")
#  legend("bottomleft", legend = c(str_c("LED n", n1, sep = " = "), str_c("    IR n", n0, sep = " = ")),
#         col = colour, lty = c(1, 2), lwd = 2, bty = "7")
}
overlap_flash_2sp <- function(x, y, col_x, col_y, rug = FALSE) {
  x0 <- obs[obs$validated_species %in% x & obs$flash == FALSE, ]$rad # - - | | - -    = FALSE
  x1 <- obs[obs$validated_species %in% x & obs$flash == TRUE, ]$rad  # sp X when flash = TRUE
  y0 <- obs[obs$validated_species %in% y & obs$flash == FALSE, ]$rad # - - | | - -    = FALSE
  y1 <- obs[obs$validated_species %in% y & obs$flash == TRUE, ]$rad # sp Y when flash = TRUE
  par(mfrow = c(2, 2)) # setting plot parameters: 2x2 windows
  overlap_flash(x, col_x, rug) # using the overlap_flash function for single species w/w/o flash-plot
  overlap_flash(y, col_y, rug) #                 - - | | - -
  overlapPlot(x0, y0,  # overlapping 2 species plot - flash = FALSE
    main = "IR", rug = rug, linecol = c(col_x, col_y),
    linetype = c(2, 2), linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey"
  )
  xy0.est <- round(overlapEst(x0, y0, type = "Dhat4"), 3)  # estimated overlap using Dhat4 - flash = FALSE
  xy1.est <- round(overlapEst(x1, y1, type = "Dhat4"), 3)  #           - - | | - -         - flash = TRUE
  legend("top", legend = str_c("Dhat4", xy0.est, sep = " = "), bty = "n") # Dhat4 label
  legend("bottomleft", legend = c(x, y), col = c(col_x, col_y), lty = c(2, 2), lwd = 2, bty = "n") # legend
  overlapPlot(x1, y1,  # overlapping 2 species plot - flash = TRUE
    main = "white LED", rug = rug, linecol = c(col_x, col_y),
    linetype = c(1, 1), linewidth = c(2, 2), olapcol = "darkgrey", extend = "lightgrey"
  )
  legend("top", legend = str_c("Dhat4", xy1.est, sep = " = "), bty = "n")
  legend("bottomleft", legend = c(x, y), col = c(col_x, col_y), lty = c(1, 1), lwd = 2, bty = "n")
  par(mfrow = c(1, 1)) # return normal plot parameters: 1x1 windows
}
obs$hour <- as.numeric(format(obs$datetime, "%H")) # setting up time cols 
obs$mins <- as.numeric(format(obs$datetime, "%M")) # in a way that overlap
obs$rad <- ((obs$hour * 60 + obs$mins)/(24 * 60)) * 2 * pi  # understands
```



```{r overlap-flash}
overlap_flash("rev", "red", rug = TRUE) 
overlap_flash_2sp("rev", "raadyr", "red", "blue", rug=TRUE)
```

Sites without white flash produces a more bumpy curve than does the sites with a flash. Proportion of foxes at sites are markedly lower before sunrise, and then higher afterwards. Could this simply be because of a lower detection rate when lacking the additional white flash? There is a resembling phenomenon happening in the evening twilight as well, right before the peak activity time of the fox, which happens before midnight.

The Dhat4 calculation reveals a larger difference in activity for foxes, than for roe deer, but seems to mainly stem from the twilight hours. Thus, it could be because of the size of the animals, rather than a reaction to the flash.

### One plot for all species

```{r overlap-flash-sp}

sp <- c("rev", "raadyr", "grevling", "elg", "hjort", "gaupe")
par(mfrow = c(2, 3))
 par(cex = 0.6)
 par(mar = c(0, 0, 0, 0), oma = c(4, 4, 0.5, 0.5))
 par(tcl = -0.25)
 par(mgp = c(2, 0.6, 0))
for (i in sp[1:3]) {
 overlap_flash(i, axes = FALSE,  ann=F, rug=T, xaxt="n")
  legend("top", legend = str_c("\n ",i), bty = "n") # art
}
overlap_flash(sp[4], main=NULL, rug=T) 
   legend("top", legend = str_c("\n ",sp[4]), bty = "n") # art 
   legend(x=-4.2, y=0.018, legend = c("IR", "white LED"), #x+y sets legend position
         col = c("black","blue"), lty = c(1, 2), lwd = 2, bty = "o")
for (i in sp[5:6]){
 overlap_flash(i, axes = FALSE,  ann=F, rug=T) 
   legend("top", legend = str_c("\n ",i), bty = "n") # art
}
 mtext("Time", side = 1, outer = TRUE, cex = 0.7, line = 2.2,
 col = "grey20")
 mtext("Density", side = 2, outer = TRUE, cex = 0.7, line = 2.2,
 col = "grey20")
```



### Controlling for day-length
```{r suntime}
# Dates <- as.POSIXct(obs$date, tz="CET")
# coords <- matrix(c(60.1, 10.6), nrow=1) # Rett nord for Oslo i longlat
# Coords <- sp::SpatialPoints(coords, proj4string=sp::CRS("+proj=longlat +datum=WGS84"))
# st <- sunTime(obs$rad, Dates, Coords)
# 
# par(mfrow=2:1)
# densityPlot(st, col='red', lwd=2, xaxt='n', main="Sun time")
# axis(1, at=c(0, 6, 12, 18, 24),
#   labels=c("midnight", "sunrise", "noon", "sunset", "midnight"))
# densityPlot(simCalls$time, lwd = 2, main = "Clock time")
# par(mfrow=c(1,1))

# TODO # Sjå density.r for oppdatert skript 
```



## Box plots
### of periods
```{r box_plots_periods}
freq     <- readRDS("freq.rds") #from line 79 in Data_exploration2_nesting.R
# Setting 20 days as an arbitrary limit
freq <- freq[freq$n.days > 19, ]
# frequency on periods, single or plural species
sp <- c("elg", "grevling","hjort", "maar", "raadyr", "rev") # Species with the most datapoints
p_freq_per <- freq %>% # plot of frequency grouped by period
  filter(validated_species %in% sp) %>%
  left_join(stations, by = "loc") %>%
  ggplot(aes(period, freq)) +
  facet_wrap(~validated_species, scales = "free", nrow = 2) +
  labs(title = "Frequencies of species in periods")

# coloured with abc-grouping
p_freq_per + geom_boxplot(outlier.shape = NA) + # Remove outliers when overlaying boxplot with original data points
  geom_jitter(aes(col = abc), width = 0.1) + coord_flip() # add points for each camera

```

### of flash on/of
```{r, box_abc_flash}
freq %>% # plot of frequency grouped by abc and flash T/F
  filter(validated_species %in% sp) %>%
  left_join(stations, by = "loc") %>%
  ggplot(aes(abc, freq)) +
  facet_wrap(~validated_species, scales = "free", nrow = 2) + # scales = "free"
  labs(title = "Frequencies of species w/ | w/o flash") +
  geom_boxplot(aes(fill = flash))
```


-----------------------------

# Some final plots I want to add in my thesis


## Maps

```{r mapsetup, message=FALSE, warning=FALSE, echo=FALSE}
v <- st_read("Basisdata_30_Viken_25833_Kommuner_GML.gml", drivers = "GML", layer = "Kommune")
o <- st_read("Basisdata_03_Oslo_25833_Kommuner_GML.gml",  drivers = "GML", layer = "Kommune")
Viken<-rbind(v,o) #legg Oslo inn i Viken
### Transform CRS to utm zone 32 (23033 for utm33)
Viken <- st_transform(Viken_sf, crs = 23032)
## automatically selected the first layer in a data source containing more than one.Reading layer `Kommune' from data source `C:\Users\Nora.Familie-PC\Documents\TorgeiR\flm\R\Basisdata_30_Viken_25833_Kommuner_GML.gml' using driver `GML'
## GDAL Message 1: 64 bit integer overflow when converting 172793172790172913172915173031
# Setting up loc-points
loc_sfg <- st_multipoint(cbind(stations$x,stations$y)) 
loc_sfc <- st_sfc(loc_sfg, crs = 23032) #23032 for utm zone 32 (23033 for utm33)
# Create data.frame with attributes
loc_df <- data.frame(loc = stations$loc,
                        cam_mod = stations$cam_mod,
                        abc = stations$abc)
# Combine data.frame and spatial data
loc_sf <- st_sf(loc_df, geometry = loc_sfc)
loc_sf
class(loc_sf)
str(loc_sf)
# loc_sf <- mutate(loc_sf, 
#                    x = purrr::map_dbl(geometry, 1), 
#                    y = purrr::map_dbl(geometry, 2))
```
```{r maps-scandi}
library(maps)
library(tmap)
Norge <- st_as_sf(map("world",'Norway', plot = FALSE, fill = TRUE))
ggplot() + geom_sf(data = Norge)
Scandinavia <- st_as_sf(map("world",c("Norway","Sweden", "Finland","Denmark","UK", "Germany"), plot = F, fill = T))

ggplot() +
  geom_sf(data =Scandinavia) +
  geom_sf(data = Viken, fill = "blue", alpha = .2) +  # Viken polygon
  annotate("rect", xmin = 7.3, xmax = 13,
           ymin = 58.5, ymax = 61.2,
           alpha = .4) +
  coord_sf(xlim = c(0,25), ylim = c(57,71)) +
  theme_bw()

# ggplot(Viken) +
#   stat_sf_coordinates()
```



```{r maps}
loc_sf <- mutate(loc_sf, 
                   x = purrr::map_dbl(geometry, 1), 
                   y = purrr::map_dbl(geometry, 2))
ggplot(data = Viken) +
        geom_sf() +     # Viken
        geom_sf_label(aes(label = navn)) + #denne kommandoen kan også hente ut namnelapper
        geom_sf(data = loc_sf, color = "red", size = 3,   # CT sites
        show.legend = "point") +
        theme_minimal() +
        theme(axis.title = element_blank()) +
        coord_sf(xlim = range(loc_sf$x), ylim = range(loc_sf$y))
#  xlim(c(190000,290000)) + ylim(c(6580000,6720000))
```



```{r}
nc <- sf::st_read(system.file("shape/nc.shp", package = "sf"), quiet = TRUE)
ggplot(nc) +
  geom_sf(aes(fill = AREA))

# If not supplied, coord_sf() will take the CRS from the first layer
# and automatically transform all other layers to use that CRS. This
# ensures that all data will correctly line up
nc_3857 <- sf::st_transform(nc, 3857)
ggplot() +
  geom_sf(data = nc) +
  geom_sf(data = nc_3857, colour = "red", fill = NA)

# Unfortunately if you plot other types of feature you'll need to use
# show.legend to tell ggplot2 what type of legend to use
nc_3857$mid <- sf::st_centroid(nc_3857$geometry)
ggplot(nc_3857) +
  geom_sf(colour = "white") +
  geom_sf(aes(geometry = mid, size = AREA), show.legend = "point")

# You can also use layers with x and y aesthetics: these are
# assumed to already be in the common CRS.
ggplot(nc) +
  geom_sf() +
  annotate("point", x = -80, y = 35, colour = "red", size = 4)

# Thanks to the power of sf, a geom_sf nicely handles varying projections
# setting the aspect ratio correctly.
library(maps)
world1 <- sf::st_as_sf(map('world', plot = FALSE, fill = TRUE))
ggplot() + geom_sf(data = world1)
map()
world2 <- sf::st_transform(
  world1,
  "+proj=laea +y_0=0 +lon_0=155 +lat_0=-90 +ellps=WGS84 +no_defs"
)
ggplot() + geom_sf(data = world2)

# To add labels, use geom_sf_label().
ggplot(nc_3857[1:3, ]) +
   geom_sf(aes(fill = AREA)) +
   geom_sf_label(aes(label = NAME))
```

```{r}
map()	# low resolution map of the world
map(wrap = c(0,360), fill = TRUE, col = 2) # pacific-centered map of the world
map(wrap = c(0, 360, NA), fill = TRUE, col = 2) # idem, without Antarctica
map('usa')	# national boundaries
map('county', 'new jersey')	# county map of New Jersey
map('state', region = c('new york', 'new jersey', 'penn'))	# map of three states
map("state", ".*dakota", myborder = 0)	# map of the dakotas
map.axes()				# show the effect of myborder = 0
if(require(mapproj))
  map('state', proj = 'bonne', param = 45)	# Bonne equal-area projection of states

# names of the San Juan islands in Washington state
map('county', 'washington,san', names = TRUE, plot = FALSE)

# national boundaries in one linetype, states in another
# (figure 5 in the reference)
map("state", interior = FALSE)
map("state", boundary = FALSE, lty = 2, add = TRUE)

# plot the ozone data on a base map
# (figure 4 in the reference)
data(ozone)
map("state", xlim = range(ozone$x), ylim = range(ozone$y))
text(ozone$x, ozone$y, ozone$median)
box()
```






### Blank images
```{r blanks, eval = FALSE}
# Filtering out timelapse-pictures in Reconyx (455 is the only Browning-camera that has significant number of pictures here)
timelapse <- obs %>% left_join(stations, by = "loc") %>%
  filter(hour %in% c(8, 13) & mins == 0  &  secs == 0  &  validated_species == "nothing") # == "nothing" to single out tinelapse-photos
summary(timelapse)
table(timelapse$hour[timelapse$cam_mod =="Reconyx"])

table(timelapse$cam_mod)


# table(timelapse$loc[timelapse$cam_mod == "Browning"])  #455 maybe has two cameras?
# table(timelapse$validated_species)  # when val_sp == "nothing" not specified: 10 other species present
# table(timelapse$validated_species[timelapse$hour == 13]) # elg:1, maar:2, menneske:3, raadyr:1
# table(timelapse$validated_species[timelapse$hour == 8]) # ekorn:1, fugl:9, gaupe:1, katt:1, menneske:1, raadyr:4, rev:2, ukjent:4

# Differentiate "nothing"-sequences and "nothing"-one-shots
# validated_species == "nothing" #by loc,
# and making an "independent event" criteria to distinguish repeated vegetation-triggering and other single-triggered events. 
# TODO
library(data.table)
setDT(obs)[, event_id := 1L + cumsum(c(0L, diff(datetime) > 720)), by=.(loc, validated_species)]

# TODO obs %>% group_by(validated_species, event_id) %>% filter(validated_species == "nothing") %>% table(obs$loc)


names(obs)
```




# Session Info
```{r sessionInfo}
sessionInfo()
# packrat
# checkpoint
```

If you want your code to be reproducible in the long-run (i.e. so you can come back to run it next month or next year), you’ll need to track the versions of the packages that your code uses.
A rigorous approach is to use _packrat_, [link](http://rstudio.github.io/packrat/), which stores packages in your project directory,
or _checkpoint_, [link](https://github.com/RevolutionAnalytics/checkpoint), which will reinstall packages available on a specified date. A quick and dirty hack is to include a chunk that runs sessionInfo() — that won’t let you easily recreate your packages as they are today, but at least you’ll know what they were.