---
title: "GLMM without locations where n.obs = 0 for every period"
author: "Torgeir Holmgard Valle"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    keep_md: true
---

```{r setup}
library(tidyverse)
library(lme4)
library(performance) # diagnostic-plots to check assumptions
library(report)      # Result-summaries in text-format
library(ggeffects)   # Estimated Marginal Means and Marginal Effects from Regression Models
                          # more at: https://strengejacke.github.io/ggeffects/
library(parameters)  # extract model-parameters etc. from (most) models
library(sjPlot)      # parameters + sjPlot probably does a similar and better job than ggeffects
library(see)         # plot-related package from the easystats-verse

# Data drom Data_exploration2_nesting.R
time.dep <- readRDS("timedep.rds")

ctrl <- c("Control_1", "Control_2", "Control_3","Control_4")
obs      <- readRDS("Observations_prepared1.rds") %>% 
  mutate(flash = ifelse(period %in% ctrl, "Control",flash))
obs <- obs %>% 
  mutate(species = validated_species,
         Hour = as.numeric(format(obs$datetime, "%H")), # for density-plots
         flash = factor(obs$flash, labels = c("IR", "LED","Control"))) # --||--
# Set global plot theme
theme_set(ggpubr::theme_classic2())
```

## Purpose

Remaking glmm_sp to test whether removing locations that never detected any individuals of each species. Bad formulation... For each species specific model, I removed the locations that never saw any individuals of said species.
And I did it to see if my models improved.

Rationale: Of course type of flash won't be able to predict any detection rate changes at cameras that never detected any individuals of a species.




### Formula
The model formula I will use is $n \sim \ time.deploy\ * flash $ for each species, and my $\alpha = 0.05$.

```{r timedep2}
sp <- c("raadyr", "rev", "hjort", "grevling", "elg", "gaupe", "ekorn", "hare", "maar")
ctrl <- c("Control_1", "Control_2", "Control_3","Control_4")
time.dep2 <- time.dep %>% 
  rename(species = validated_species) %>%  #shortening name
#  filter(species %in% sp) %>% #filtering out species
  # including Control as part of the flash-column, since it differs from flash=0
  mutate(flash = factor(
        ifelse(period %in% ctrl, "Control", flash)),
        week = lubridate::isoweek(date),
        period = factor(period))
time.dep2 <- time.dep2 %>% 
   mutate(flash = fct_relevel(flash, "Control","0","1")) # relevel to make Control the model intercept
levels(time.dep2$flash) <- c("Control", "IR", "LED")
levels(time.dep2$period) <- c("IR_1", "IR_2", "LED_1", "LED_2", "Control_1", "Control_2", "Control_3", "Control_4")
```



```{r period-median, message=FALSE}
# find median period length
time.period <- time.dep2 %>% group_by(loc, period, flash) %>% 
  summarise(period_length = max(time.deploy))

# checking shortest periods
time.period %>% arrange(period_length) # 1 period (LED) is 0 days

# then merge lengths and filter out period of 0 days
time.dep3 <- time.dep2 %>% left_join(time.period) %>% 
  filter(period_length > 0)

# find median length 
time.period %>% filter(flash == "LED") %>%  
  summary() # median period length 85 days, mean: 84
time.period %>% filter(flash == "IR") %>%  
  summary() # median period length 79 days, mean: 89

# extract lengths of each unique period
h <- time.dep3 %>% group_by(loc, period, period_length, flash)%>% nest() %>% 
  select(!data) 
#extracting median and multiplying by 10, to use in the correctly scaled plot
hh <-       h$period_length[h$flash == "LED"] %>%  median()       # median white LED
hh <- c(hh, h$period_length[h$flash == "IR" ] %>%  median()) * 10 # + median IR
# smallest median 
h <- min(hh)
```




_Lastly, performing the filter:_
```{r time.dep4-filtr}
habitat <- readRDS("habitat.rds") %>% 
  mutate(loc = factor(loc))
# filtering out periods longer than (shortest) median length.
time.dep4 <- time.dep3 %>% filter(time.deploy < h/10) %>%  # h is normal scale, must be rescaled by /10
  left_join(habitat, by = "loc") 
# time.dep5 <- time.dep4 %>% filter(loc[n.obs > 0])

```


# Modelling


## Roe deer


```{r sp, eval=FALSE}
# sp ="maar"
loc1 <- time.dep4$loc[time.dep4$n.obs > 0 & time.dep4$species %in% sp] %>% unique()
time_sp <- filter(time.dep4, species %in% sp, loc %in% loc1) #.dep4 = trimmed data
# time_sp <- filter(time.dep5, species %in% sp, !period %in% ctrl) #.dep5 = test-transformations
# Model
m_sp  <- lme4::glmer(n.obs ~ time.deploy * flash + # fixed effects
            (1 | loc) + (1 | week) + (1|habitat), # random effects
            data   = time_sp,
            family = poisson) # poisson family of distributions
# ggeffect calls effects::Effect
p_sp    <- ggeffects::ggeffect(m_sp, terms = c("time.deploy [all]", "flash"))
# Diagnostics
assumpt <- performance::check_model(m_sp) # check assumptions
# N sites
length(loc1)
```



```{r sp-report, eval=FALSE}
# Summary, report, model
summary(m_sp)
r_sp <- report::report(m_sp) # text-summary of my model, to include in a report
para_sp  <- model_parameters(m_sp,   standardize = "refit", two_sd = TRUE, exponentiate = TRUE)
# saveRDS(m_sp, file = paste0("m2_",sp,".rds")) # save model objects as shortcut for when editing etc.
```



`r sp = "raadyr"`
```{r raadyr, ref.label='sp'}
# filter species
#<<sp>>
```




```{r raadyr-report, ref.label='sp-report'}
#<<sp-report>>

assumpt
```


## Roe deer plot

```{r, eval=FALSE}
# just a shortcut for editing
sp="raadyr"
m_sp <- readRDS("m_raadyr.rds")
p_sp <- ggeffects::ggeffect(m_sp, terms = c("time.deploy [all]", "flash"))
```


```{r parameters}
library(cowplot) # to make grid-plots
library(magick)
# ggpredict
p_sp1 <- plot(p_sp, ci.style = c("dash"), line.size = 1, #ci.styles: “ribbon”, “errorbar”, “dash”, “dot”
               colors = c("black","#e41a1c","#377eb8"),
              ) +
   labs(title="", x="Time since deployment (per 10 days) \ ", y="Detection rate") +
   ggpubr::theme_classic2() +
  theme(legend.position = "bottom", legend.title = element_blank(),
        axis.title = element_text(size = 11)) 

# standardized plots aren't very different, other than on the scale
#plot(para_sp, size_text = 3) + labs(title = paste0(sp, " GLMM parameters") ,
#                                    subtitle = 'standardize  = "refit" ')

# Equivalence test
result <- equivalence_test(m_sp)
result

# labels for equivalence test - prettier to the human eye
par_lab <- c("Time", "IR", "LED", "Time * IR", "Time * LED")
par_lab <- par_lab[5:1]
# Equivalence plot
p_eq <- plot(result) + labs(y = "Log-Mean") + 
    scale_x_discrete(labels = par_lab) + # new axis names
    guides(colour = guide_legend(nrow = 2, override.aes = list(size = .5),
                                 title.theme = element_text( 
                                   size=10, #adjusting legend appearance
                                   face="italic"))) 

# Density plots
p_dens <- obs %>% filter(species %in% sp) %>% 
  mutate(flash = fct_shift(flash,-1)) %>% #reordering flash-factor
  ggplot(aes(Hour)) +
  geom_bar(col="black", fill="white") +
  geom_density(aes(y=..density..*20*count, #scaling density with the count
                   fill=flash, alpha=.1),
               show.legend = c(alpha = F), bw=1.2) +
  scale_x_continuous(breaks = seq(0,23, by=4)) + # which x-ticks
  scale_y_continuous(n.breaks = 6) + # n y-ticks
  theme(legend.position = c(1, 1), legend.justification = c(.1, 2), #legend placement
        legend.title = element_blank(), legend.key.size = unit(2, 'mm'))+ #size
  scale_fill_bluebrown(reverse = T) 


# cowplot::plot_grid(NULL,p_eq,p_dens,NULL,
#                    #nrow = 2,
#                    # rel_widths = c(3,4),
#                    rel_heights = c(3,2),
#                    labels="auto"
#                    #align = "h"
# )


sp_file <- paste0("jpg/",sp,".JPG")
jpg <- ggdraw() + draw_image(sp_file)

p_grid <- cowplot::plot_grid(p_sp1,
                             p_eq,
                             p_dens,
                             jpg,
                   #nrow = 2,
                   rel_widths = c(3,4),
                   rel_heights = c(3,2),
                   labels="auto"
                   #align = "h"
) 
p_grid #+ draw_image(sp_file, scale = .4, x = 0.9,
       #  hjust = 1, halign = 1, valign = 0)
```




```{r sp-report2}
summary(r_sp)
as.report_table(r_sp)
```


```{r raadyr-objects}
# Storing species-specific objects for later
# Model
m_raa    = m_sp
# ggpredict 
p_raa    = p_sp
# report-object
r_raa    = r_sp
# parameters refit
para_raa = para_sp
```


sp
sp-report
sp-report2
parameters

objects


#### Skrivestopp

-------------------------------------------------------------

## Red Fox

`r sp = "rev"`

```{r rev, ref.label=c('sp','sp-report') }
```
### Plot

```{r rev2, ref.label=c('sp-report2','parameters')}
```


```{r rev-objects }
# Model
m_rev    = m_sp
# ggpredict 
p_rev    = p_sp
# report-object
r_rev    = r_sp
# parameters refit
para_rev = para_sp
```


```{r knit-exit, eval=TRUE
}
knitr::knit_exit() # to exit knitting process here instead of at the document end
```


## Badger

`r sp = "grevling"`

```{r grevling, ref.label=c('sp','sp-report') }
```

### Plot

```{r grevling2, ref.label=c('sp-report2','parameters')}
```


```{r grevling-objects }
# Model
m_grvl    = m_sp
# ggpredict 
p_grvl    = p_sp
# report-object
r_grvl    = r_sp
# parameters refit
para_grvl = para_sp
```


## Moose

`r sp = "elg"`

```{r elg, ref.label=c('sp','sp-report') }
```

### Plot

```{r elg2, ref.label=c('sp-report2','parameters')}
```


```{r elg-objects }
# Model
m_elg    = m_sp
# ggpredict 
p_elg    = p_sp
# report-object
r_elg    = r_sp
# parameters refit
para_elg = para_sp
```

## Red deer

`r sp = "hjort"`

```{r hjort, ref.label=c('sp','sp-report') }
```

### Plot

```{r hjort2, ref.label=c('sp-report2','parameters')}
```


```{r hjort-objects }
# Model
m_hjort    = m_sp
# ggpredict 
p_hjort    = p_sp
# report-object
r_hjort    = r_sp
# parameters refit
para_hjort = para_sp
```


## Lynx

`r sp = "gaupe"`

```{r gaupe, ref.label=c('sp','sp-report') }
```

### Plot

```{r gaupe2, ref.label=c('sp-report2','parameters')}
```


```{r gaupe-objects }
# Model
m_gaup    = m_sp
# ggpredict 
p_gaup    = p_sp
# report-object
r_gaup    = r_sp
# parameters refit
para_gaup = para_sp
```


# Small species

Added late, because I was unsure about whether it made sense to include them or not.
After having learned about the random effects, I think it does make sense, even though the cameras in my study were set up with the original aim of photo capturing lynx.

## Hare

`r sp = "hare"`

```{r hare, ref.label=c('sp','sp-report') }
```

### Plot

```{r hare2, ref.label=c('sp-report2','parameters')}
```


```{r hare-objects }
# Model
m_hare    = m_sp
# ggpredict 
p_hare    = p_sp
# report-object
r_hare    = r_sp
# parameters refit
para_hare = para_sp
```


## Red squirrel

`r sp = "ekorn"`

```{r ekorn, ref.label=c('sp','sp-report') }
```

### Plot

```{r ekorn2, ref.label=c('sp-report2','parameters')}
```


```{r ekorn-objects }
# Model
m_ekorn    = m_sp
# ggpredict 
p_ekorn    = p_sp
# report-object
r_ekorn    = r_sp
# parameters refit
para_ekorn = para_sp
```



## European Pine marten

`r sp = "maar"`

```{r maar, ref.label=c('sp','sp-report') }
```

### Plot

```{r maar2, ref.label=c('sp-report2','parameters')}
```


```{r maar-objects }
# Model
m_maar    = m_sp
# ggpredict 
p_maar    = p_sp
# report-object
r_maar    = r_sp
# parameters refit
para_maar = para_sp
```



----------------------------------------------

# All models


## Joint forest-plots

Plots divided into size-based groups of three x three species.
Thus, ungulates are one group, the largest carnivores are one, and the smallest three form the most mixed group of hare, squirrel and pine marten 

```{r mod-plot}
p_mds1<- sjPlot::plot_models(m_raa, m_rev, m_grvl, m_elg, m_hjort, spacing = .8,
                    legend.title = "Species",
                    m.labels = c("Roe deer","Red fox","Badger","Moose","Red deer"))
p_mds2 <- sjPlot::plot_models(m_hare, m_maar,m_ekorn,m_gaup, spacing = .8,
                    legend.title = "Species",
                    m.labels = c("Hare","Pine marten","Red squirrel","Lynx"))
p_mds1;p_mds2
# saveRDS(p_mds1, "pmds1.rds")
# saveRDS(p_mds2, "pmds2.rds")
```


## Model comparison-plots

Whichs species' detection rates was best explained by the my model formula?

```{r mod-comparisons}
m_compare <- compare_performance(m_raa,m_rev,m_grvl,m_elg,m_hjort,
                                 m_gaup, m_hare, m_ekorn, m_maar,
            metrics = "common", rank=T) # "common" will compute AIC, BIC, R2, ICC and RMSE
m_compare
m_compare %>% plot() #A `range` must be provided for data with only one observation.
# test_performance(m_raa,m_rev,m_grvl,m_elg,m_hjort,m_gaup) #models don't have the same response variable, 
# which is because they are different subsets of each other
```


```{r mod-checks, eval=FALSE}
check_overdispersion(m_raa)   # No overdispersion detected
check_overdispersion(m_rev)   # -     -    | |    -    -
check_overdispersion(m_grvl)  # -     -    | |    -    -
check_overdispersion(m_elg)   # -     -    | |    -    -
check_overdispersion(m_hjort) # -     -    | |    -    -
check_overdispersion(m_gaup)  # -     -    | |    -    -
check_zeroinflation(m_raa)   # Model seems ok, ratio of observed and 
check_zeroinflation(m_rev)   #    predicted zeros is within the tolerance range.  
check_zeroinflation(m_grvl)  #          -     -    | |    -    -   
check_zeroinflation(m_elg)   #          -     -    | |    -    -   
check_zeroinflation(m_hjort) #          -     -    | |    -    -  
check_zeroinflation(m_gaup)  #          -     -    | |    -    -
check_singularity(m_raa)   # FALSE
check_singularity(m_rev)   #--||--
check_singularity(m_grvl)  #--||--   
check_singularity(m_elg)   #--||--   
check_singularity(m_hjort) #--||--  
check_singularity(m_gaup)  #--||--
```
















--------------------------------------------


### Blueprint for other species in chunk below:

<!--`r sp = "xx"`-->

```{r xx, ref.label=c('sp','sp-report'), eval=FALSE}
```
```{r xx2, ref.label=c('sp-report2','parameters'), eval=FALSE}
```


```{r xx-objects, eval=FALSE}
# Model
m_xx    = m_sp
# ggpredict 
p_xx    = p_sp
# report-object
r_xx    = r_sp
# parameters refit
para_xx = para_sp
```


----------------------------------------------

# SessionInfo

```{r sessionInfo}
sessionInfo()

report_parameters(sessionInfo()) # output to include in Appendix
```

